---
title: "Tidy Models template"
author: "Peter Boshe"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Context

We are going to model the data in a way that we are able to predict the count of bottles

## load clean data

```{r}
library(tidymodels)
library(tidyverse)
library(doMC)
library(poissonreg)
tidymodels_prefer(quiet = TRUE)

# parameters

   #input
file_raw <- here::here("docs/data/clean_waste_base_data.rds")

# set theme

theme_set(theme_bw())

```

## E.D.A

```{r}
waste_df <- read_rds(file_raw)



#check percent of bottlecount is 0
# mean(waste_df$bottle_count < 1)

# relation between amount and bottle count

# correlation
# cor(waste_df$amount, waste_df$bottle_count)

# table(waste_df$product_category, waste_df$scan_country)
# table(waste_df$product_category, waste_df$manufacturer_country)


## Cutting waste_df$amount into waste_df$amount_rec
waste_df$amount_rec <- cut(waste_df$amount,
  include.lowest = TRUE,
  right = FALSE,
  dig.lab = 4,
  breaks = c(0, 500, 1000, 2000, 5000, 20000)
)

glimpse(waste_df)

# table(waste_df$product_category, waste_df$amount_rec)


# plot histogram

waste_df |> 
  ggplot(aes(bottle_count)) + 
  geom_histogram() +
  scale_x_log10()


# bar plot of product category and bottlecount
waste_df |>
  group_by(product_category) |> 
  arrange(desc(bottle_count)) |> 
  ungroup() |> 
ggplot(aes(bottle_count, fct_rev(fct_infreq(product_category)),  fill = amount_rec)) +
  geom_col()

```

-   we have had to transform the bottle count variable by log base of 10 since the distribution is skewed
-   we have no zero counts due to our cleaning pipeline
-   we are goint to uses a regular poisson regression, as opposed to the zero inflated poisson (ZIP) version which is a mixed effect model

however let's review the definition of poisson distribution as per the oracle(wikipedia)

Poisson distribution

:   In probability theory and statistics, the Poisson distribution is a discrete probability distribution that expresses the probability of a given number of events occurring in a fixed interval of time or space if these events occur with a known constant mean rate and independently of the time since the last event.

Our data is of time series nature, I am not sure my previous decision to go with poisson distribution is justified.

hmm... but then again `batch` is a categorical variable and can be treated as such

**approach 1**

we are going to first neglect the effect of the time series

## Data allocation

```{r}
waste_df <- waste_df |>  
  mutate(batch = as.numeric(batch)) |> 
  arrange(batch) %>% 
  mutate(.row = row_number()) # to simplify indexing
waste_df


# inspect the proportion of the split if we split by  batch
split_prop <- (last(which(waste_df$batch <= 13))) / nrow(waste_df)

indices <-
  list(analysis   = waste_df$.row[waste_df$batch <= 13], 
       assessment = waste_df$.row[waste_df$batch >  13]
  )

waste_split <- make_splits(indices, waste_df %>% select(-.row))
waste_train <- training(waste_split)
waste_test <- testing(waste_split)


# folds for cross validation

waste_folds <- vfold_cv(waste_train,v = 20)

```

## feature engineering recipes

-   categorical data will need to be encoded for the poisson regression method

```{r}

# recipe for poisson reg
waste_rec <- recipe(bottle_count ~ batch + product_category + brand_name + manufacturer_country + scan_country + amount + units + amount_rec, data = waste_train) |> 
  step_novel(all_nominal_predictors()) |> 
  step_dummy(all_nominal_predictors()) |> 
  step_nzv(all_predictors()) |> 
  step_normalize(all_numeric_predictors())

# simple recipe for random forest
model_vars <- workflow_variables(outcomes = bottle_count,
                                 predictors = everything())  

```

## calling all models

```{r}

poisson_reg_glm_spec <-
  poisson_reg() %>%
  set_engine('glm')

rand_forest_ranger_spec <-
  rand_forest(mtry = tune(), min_n = tune()) %>%
  set_engine('ranger') %>%
  set_mode('regression')



```

## setting our workflows

```{r}
poisson_wf <- workflow_set(list(basic = waste_rec), list(poisson = poisson_reg_glm_spec))

randforest_wf <- workflow_set(list(simple = model_vars), list(randforest = rand_forest_ranger_spec))

# combining our workflows
all_wflows <- 
  bind_rows(poisson_wf, randforest_wf)
all_wflows
```

## tuning our parameters

```{r,eval=TRUE}

ctrl <- control_grid(save_pred = TRUE,
                     save_workflow = TRUE)

doMC::registerDoMC(cores = 8)

full_res_time <-  system.time(
  grid_results <- 
  all_wflows |> 
  workflow_map(
    "tune_grid",
    seed = 123,
    verbose = TRUE,
    grid = 25,
    resamples = waste_folds,
    control = ctrl
  )
)
registerDoSEQ()
BRRR::skrrrahh(44)

# takes 6 and a half minutes

```

## evaluating our models

```{r}

# metrics <- collect_metrics(grid_results)

(autoplot(
   grid_results,
   rank_metric = "rmse",  
   metric = "rmse",       
   select_best = TRUE    
))
(autoplot(
   grid_results,
   rank_metric = "rsq",  
   metric = "rsq",       
   select_best = TRUE    
))
```

```{r}


# function does not work
grid_results |> show_best(metric = "rmse")

```
