---
title: "Waste data exploration"
author: "Peter Boshe"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    fig_caption: true
    theme: flatly
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = 'center', collapse = TRUE, echo = FALSE, message = FALSE)
```


```{r}
# Packages
library(tidyverse)
library(gt)
library(countrycode)
library(plotly)

# Parameters
  #input
file_raw <- here::here("docs/data/clean_waste_base_data.rds")
file_flags <- here::here("docs/data-raw/Country_Flags.csv")

# theme update

theme_set(theme_classic())

```


Lets explore our data

```{r}
df <- read_rds(file_raw)



df |> 
  glimpse()
```


- batch amount progression
```{r}
df |> 
  group_by(batch) |> 
  summarise(total_bottles = sum(bottle_count)) |> 
  ggplot(aes(batch,total_bottles)) +
  geom_col() +
  geom_line(aes(group = 1)) +
  geom_point() 
  # geom_label(aes(label = total_bottles))



```



- product category distribution

```{r}
df |> 
  group_by(product_category) |> 
  summarise(total = sum(bottle_count)) |> 
  arrange(total) |> 
  ggplot(aes(fct_inorder(product_category), total)) +
  geom_col() +
  coord_flip()

```

- product label distribution

```{r}
df |> 
  mutate(product_label = as_factor(product_label)) |> 
  group_by(product_label,brand_name) |> 
  summarise(total = sum(bottle_count)) |>
  arrange(desc(total)) |> #view()
  filter(total > 5000) |> 
  ggplot(aes(droplevels(product_label), total, group = brand_name)) + 
  geom_col() +
  coord_flip() +
  theme(
    axis.text.x = element_text(angle = 90)
  )

# REVIEW figure out why you cant order the caategories correctly
  
```


- units distribution

```{r}
df |> 
  count(units) |> 
  ggplot(aes(units, n)) + 
  geom_col()
```

- scan country distribution

```{r}
df |> 
  group_by(scan_country) |> 
  summarise(total = sum(bottle_count)) |> 
  arrange(total) |> 
  ggplot(aes(total, fct_inorder(scan_country))) + 
  geom_col() +
  geom_label(aes(label = total))
```



- manufacturer country distribution

```{r}
df |> 
  group_by(manufacturer_country) |> 
  summarise(total = sum(bottle_count)) |> 
  slice_max(order_by = total, prop = .2) |> 
  arrange(total) |> 
  ggplot(aes(total, fct_inorder(droplevels(manufacturer_country)))) + 
  geom_col() +
  geom_label(aes(label = total))

# HACK great way to slice and reorder
# REVIEW to checkways to summarize the slice better

```



- brand name distribution and scan country

```{r}
# df |> 
#   group_by(brand_name, scan_country) |> 
#   summarise(total = sum(bottle_count)) |> 
#   group_by(scan_country) |> 
#   slice_max(order_by = total,n = 5) |> 
#   arrange(total) |>
#   # fct_reorder(df$brand_name,total, min, .desc = TRUE)
#   # na.omit() |> 
#   ggplot(aes(total, fct_inorder(droplevels(brand_name)))) +
#   geom_col() +
#   facet_wrap( ~ scan_country, scales = "free")

df |> 
    group_by(brand_name, scan_country) |> 
    summarise(total = sum(bottle_count)) |> 
    group_by(scan_country) |> mutate(scan_country=fct_reorder(scan_country, total, .desc = TRUE))|> 
    slice_max(order_by = total,n = 5) |> 
    arrange(total) -> countries_arranged

## Recoding countries_arranged$scan_country into countries_arranged$scan_country_rec
countries_arranged$scan_country_rec <- countries_arranged$scan_country %>%
  fct_recode(
    "Angola" = "ao",
    "Brazil" = "br",
    "Canada" = "ca",
    "Switzerland" = "ch",
    "Spain" = "es",
    "France" = "fr",
    "Great Britain" = "gb",
    "Kenya" = "ke",
    "Malawi" = "mw",
    "Mozambique" = "mz",
    "Netherlands" = "nl",
    "Rwanda" = "rw",
    "Tanzania" = "tz",
    "Uganda" = "ug",
    "South Africa" = "za",
    "Zambia" = "zm"
  )

    # fct_reorder(df$brand_name,total, min, .desc = TRUE)
    # na.omit() |> 
    ggplot(countries_arranged,aes(total, fct_inorder(droplevels(brand_name)))) +
    geom_col() +
    facet_wrap( ~ fct_rev(fct_inorder(scan_country_rec)), scales = "free")

```



- year and month distribution


```{r}
df |> 
  group_by(year, month, product_category) |> 
  summarise(total_bottles = sum(bottle_count)) |> 
  ggplot(aes(month,total_bottles, fill = product_category)) +
  geom_col(col = "white") +
  facet_wrap( ~ year)


df |> 
  group_by(year, month, product_category) |> 
  summarise(total_bottles = sum(bottle_count)) |> 
  ggplot(aes(month,total_bottles)) +
  geom_boxplot() +
  facet_wrap( ~ year)

```

# investigating outliers

```{r}
outliers <- df |> 
  nest(data = -c(batch,brand_name, scan_country)) |> 
  mutate(total_bottles = map_dbl(data, ~ sum(.$bottle_count))) |> 
  unnest(cols = c(data)) |> 
  ggplot(aes(batch, total_bottles)) +
  geom_jitter(aes(fill = product_category,
                  shape = scan_country,
                  group = scan_country,
                  size = bottle_count),
              alpha = 0.4,
              width = 0.8) +
  geom_boxplot(alpha = 0.5,
               outlier.shape = NA,
               outlier.alpha = 0.1,
               size = 2)
ggplotly(outliers,
         tooltip = c("size","y", "fill", "shape"))

# TODO add median line
```







- distribution of amounts

```{r}
df |> 
  filter(units == "ml") |>
  mutate(amount = round(amount),
         amount_category = fct_inorder(case_when(
           amount < 500 ~ "less than 500",
           amount >= 500 & amount < 1000 ~ "500-1000",
           amount >= 1000 & amount < 2000 ~ "1000-2000",
           amount >= 2000 & amount < 5000 ~ "2000-5000",
           TRUE ~ "greater than 5000"
           ))
  ) |> 
  mutate(amount_category = fct_relevel(amount_category, "greater than 5000", after = 4)) |> # HACK releveling default levels
  count(amount_category) |> 
  ggplot(aes(n, fct_rev(amount_category))) + # HACK flipping your levels with forcats
  geom_col()
```



# geospatial map

```{r}
library(leaflet)
library(rnaturalearth)
library(sp)
map <- ne_countries()
# names(map)[names(map) == "iso_a2"] <- "ISO2"
# names(map)[names(map) == "name"] <- "NAME"

# summary(map)
# names(map)
# class(map)
# str(map)
# map$iso_a3
# map$iso_a2
# map$name

# plot(map)


scan_country_count <- df |> 
  group_by(scan_country) |> 
  summarise(total_bottles=sum(bottle_count))
scan_country_count


map$bottle_count <- as_vector(scan_country_count[match(map$iso_a2,str_to_upper(scan_country_count$scan_country)), "total_bottles"])
```



```{r}
library(DT)
DT::datatable(map@data[, c("iso_a2", "name", "bottle_count")],
  rownames = FALSE, options = list(pageLength = 10)
)
```

# plotting the map

```{r}

pal <- colorBin(
  palette = "viridis", domain = map$bottle_count,
  bins = seq(0, max(map$bottle_count, na.rm = TRUE) + 10, by = 10)
)


map$labels <- paste0(
  "<strong> Country: </strong> ",
  map$name, "<br/> ",
  "<strong> Bottle Count: </strong> ",
  map$bottle_count, "<br/> "
) %>%
  lapply(htmltools::HTML)

leaflet(map) |> 
  addTiles() |> 
  setView(lng = 25, lat = 0, zoom = 4) %>%
  addPolygons(
    fillColor = ~ pal(bottle_count),
    color = "white",
    fillOpacity = 0.7,
    label = ~labels,
    highlight = highlightOptions(
      color = "black",
      bringToFront = TRUE
    )
  ) 
  # leaflet::addLegend(
  #   pal = pal ,values = ~bottle_count,
  #   opacity = 0.7, title = "Bottle Count"
  # )
  # 



```



# batch progression faceted by scan countries

```{r}
library(gghighlight)

df |>
  group_by(scan_country, batch) |>
  summarise(total_bottles = sum(bottle_count),
            .groups = "drop") |>
  nest(data = -scan_country) |>
  mutate(scan_country_total = map_dbl(data, ~ sum(.$total_bottles))) |>
  unnest(cols = data) |> 
  ggplot(aes(batch,
             total_bottles, 
             group = scan_country,
             colour = scan_country)) +
  # geom_col() +
  geom_line() +
  geom_point() +
  gghighlight(scan_country_total > 5000) +
  # facet_wrap(~ scan_country,
  #            ncol = 2,
  #            scales = "free_y")+
  theme(
    strip.background = element_blank()
  )




```



```{r}
df |>
  group_by(scan_country, batch) |>
  summarise(total_bottles = sum(bottle_count),
            .groups = "drop") |>
  nest(data = -scan_country) |>
  mutate(scan_country_total = map_dbl(data, ~ sum(.$total_bottles))) |>
  arrange(desc(scan_country_total)) |>
  mutate(scan_country= fct_inorder(scan_country)) |> 
  unnest(cols = data) |> 
  filter(scan_country_total > 5000) |>
  ggplot(aes(batch,
             total_bottles, 
             group = scan_country,
             colour = scan_country)) +
  # geom_col() +
  geom_line() +
  geom_point() +
  facet_wrap(~ scan_country,
             ncol = 1,
             scales = "free_y")+
  theme(
    strip.background = element_blank()
  )

  
```

# Table of values

```{r}


flags <- read_csv(file_flags) |> 
  mutate(Code = as_factor(str_to_lower(countrycode(sourcevar = Country, origin = "country.name", destination = "iso2c", warn = FALSE)))) |> 
  select(Code,
         flag_URL = ImageURL)


# df <- df |> 
#   left_join(flags, by = c("scan_country"="Code"))

df |>
  group_by(scan_country, batch) |>
  summarise(total_bottles = sum(bottle_count),
            .groups = "drop") |>
  nest(data = -scan_country) |>
  mutate(scan_country_total = map_dbl(data, ~ sum(.$total_bottles))) |>
  arrange(desc(scan_country_total)) |>
  mutate(scan_country= fct_inorder(scan_country)) |> 
  unnest(cols = data) -> table

table$scan_country_rec <- table$scan_country %>%
  fct_recode(
    "Angola" = "ao",
    "Brazil" = "br",
    "Canada" = "ca",
    "Switzerland" = "ch",
    "Spain" = "es",
    "France" = "fr",
    "Great Britain" = "gb",
    "Kenya" = "ke",
    "Malawi" = "mw",
    "Mozambique" = "mz",
    "Netherlands" = "nl",
    "Rwanda" = "rw",
    "Tanzania" = "tz",
    "Uganda" = "ug",
    "South Africa" = "za",
    "Zambia" = "zm"
  )

 table |> 
  left_join(flags, by = c("scan_country" = "Code")) |>
  nest(data = -scan_country_rec) |>
  mutate(last_change = map_dbl(data, ~ . |>  # HACK how to map a piping function
    mutate(change = c(0, diff(total_bottles))) |>
    slice_tail(n = 1) |>
    pull(change)),
    last_month_rec = map_dbl(data, ~. |> 
                               slice_tail(n = 1) |> 
                               pull(total_bottles))) |>
  unnest() |>
  distinct(flag_URL,
           scan_country,
           scan_country_rec, 
           last_month_rec,
           last_change,
           scan_country_total) |> 
  select(flag_URL,
         scan_country_rec, 
           last_month_rec,
           last_change,
           scan_country_total) |> 
  gt() |> 
  cols_label(flag_URL = "",
             scan_country_rec = md("**Country**"),
             last_month_rec = md("**Latest Count**"),
             last_change = md("**Difference<br>from previous<br>month**"),
             scan_country_total = md("**Total bottles<br>scanned**"))|> 
    gt::text_transform(
    #Apply a function to a column
    locations = cells_body(c(flag_URL)),
    fn = function(x) {
      #Return an image of set dimensions
      web_image(
        url = x,
        height = 12
      )
    }) |> 
  cols_width(c(flag_URL) ~ px(30)) |> 
   tab_header(title = md("**Summary Table**")) |> 
   opt_align_table_header(align = "left") |> 
   tab_style(
     locations = cells_title(groups = "title"),
     style = list(
       cell_text(weight = "bold",
                 size = 24)
     )
   ) |> 
   fmt_number(columns = 3:5,
              sep_mark = ",",
              decimals = 0) |> 
  data_color(
    columns = 5,
    colors = scales::col_numeric(
      palette = as.character(paletteer::paletteer_d(palette = "ggsci::red_material")),
      domain = NULL
    )
  )
  





```





